Parsing Postfix log files

### 2009-04-04 doc/logparser-implementation.tex
Automatically generate the list of graphs, tables, and other included files.
Added graph showing average number of rules tried per program. Wrote a brief comparison between inserting and not inserting results. Removed the explanation of why mail loops occur, it's not important. Improved the blurb about the machine used for generating statistics.

There is no correlation between how often an action is specified by rules
and how often an action is invoked because a rule recognises a log line.
\Graphref{Number of times each action was invoked excluding 22 and 62--68}
shows the number of times each action was invoked when parsing the
\numberOFlogFILES{} log files used to generate the statistics in
\sectionref{parser efficiency}, \textit{excluding\/} log files 22 and
62--68, because their content is extremely skewed by two mail loops (see
\sectionref{Explaining the peaks in log file size}).  The action most
commonly specified by rules, \action{DELIVERY\_REJECTED}, is the third most
commonly invoked action; the most commonly invoked actions,
\action{CONNECT} and \action{DISCONNECT}, are each specified by only one
rule.  The \action{UNINTERESTING} action, the third most commonly specified
action, is halfway down the graph.

@@

% do not reformat this!
\begin{table}[ht]

    \caption{Example of a rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}

\end{table}

>>>>

There is no correlation between how often an action is specified by rules
and how often an action is invoked because a rule recognises a log line.
\Graphref{Number of times each action was invoked excluding 22 and 62--68}
shows the number of times each action was invoked when parsing the
\numberOFlogFILES{} log files used to generate the statistics in
\sectionref{parser efficiency}, \textit{excluding\/} log files 22 and
62--68, because their content is extremely skewed by two mail loops.  The
action most commonly specified by rules, \action{DELIVERY\_REJECTED}, is
the third most commonly invoked action; the most commonly invoked actions,
\action{CONNECT} and \action{DISCONNECT}, are each specified by only one
rule.  The \action{UNINTERESTING} action, the third most commonly specified
action, is halfway down the graph.

@@

% do not reformat this!
\begin{table}[ht]
    \caption{Example of a rule used by \parsernameshort{}}
    \empty{}\label{Example rule in implementation table}
    \begin{tabular}{ll}
        \tabletopline{}%
        \textbf{Field}      & \textbf{Value}                                    \\
        \tablemiddleline{}%
        name                & Unknown sender domain                             \\
        description         & We do not accept mail from unknown domains        \\
        restriction\_name   & reject\_unknown\_sender\_domain                   \\
        program             & \daemon{smtpd}                                    \\
        regex               & \verb!^__RESTRICTION_START__ <(__SENDER__)>: !    \\
                            & \verb!Sender address rejected: Domain not found;! \\
                            & \verb!from=<\k<sender>> to=<(__RECIPIENT__)> !    \\
                            & \verb!proto=E?SMTP helo=<(__HELO__)>$!            \\
        result\_data        &                                                   \\
        connection\_data    &                                                   \\
        action              & \action{DELIVERY\_REJECTED}                       \\
        hits                & 0                                                 \\
        hits\_total         & 0                                                 \\
        priority            & 0                                                 \\
        debug               & 0                                                 \\
        %cluster\_group      & 400                                               \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

### 2009-04-04 doc/logparser-results.tex
Added graph showing average number of rules tried per program. Wrote a brief comparison between inserting and not inserting results. Removed the explanation of why mail loops occur, it's not important. Improved the blurb about the machine used for generating statistics.

XXX DO A TEST-SOME RUN AND SAVE THE RESULTS TO THE DATABASE, THEN COMPARE
THE TIME WITH A RUN THAT DOES NOT SAVE RESULTS\@; PUT THE COMPARISON IN THE
INTRODUCTION, NEAR THE MACHINE SPECS\@.

XXX THINK ABOUT THE ORDERING OF TABLES AND GRAPHS\@; MAYBE REARRANGE THEM
SO THAT THEY APPEAR IN THE ORDER THEY ARE REFERENCED\@?

XXX SHOULD I USE $\backslash{}$clearpage{} OR SOMETHING TO AFFECT
PLACEMENT\@?  GRAPHS AND TABLES TEND TO MOVE QUITE FAR IN SOME OF THE
SUBSECTIONS, THOUGH MAKING THE TEXT LONGER MIGHT HELP TOO\@.  IDEALLY I
WANT TO HAVE TEXT, THEN ALL GRAPHS AND TABLES, THEN THE NEXT SECTION\@; I
DO NOT LIKE THAT GRAPHS/TABLES SOMETIMES APPEAR BEFORE THEY ARE
REFERENCED\@.  I MIGHT NEED TO REDUCE THE HEIGHT OF THE GRAPHS TOO\@.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} shows that, as expected, far more mails
are received on weekdays than at weekends.  The mail server in question is
a production mail server handling mail for a university department; the
benefit of using this server is that its log files exhibit the
idiosyncrasies and peculiarities a mail server in the wild must deal with,
but the downside is that significantly altering the configuration to
accommodate this project is not an option.  Median log file size is 50MB,
containing 285,000 log lines; large scale mail servers would have much
larger log files.  Note that \graphref{Mails received per day} and
\tableref{Number of mails received per day: statistics} show the number of
mails received by \acronym{SMTP} only; in particular the mail loops
noticeable in later graphs do not contribute to these figures.

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, and the parsing times from the 10 runs averaged.  Saving
results to the database was disabled for the test runs, because that
dominates the run time of the program, and the tests are aimed at measuring
the speed of \parsername{} rather than the speed of the database and the
disks the database is stored on.  The computer used for test runs was a
Dell Optiplex 745, shown in \tableref{Details of the computer used to
generate statistics}.

@@

Parsing all \numberOFlogFILES{} log files in one run took
\input{build/include-timing-run-duration}.  The computer was dedicated to
the task of gathering statistics from test runs, and was not used for any
other purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly.  The
ratios increase (i.e.\ improve) for log files 22 and 62--68 despite their
larger than average size (shown in \graphref{parsing time vs file size vs
number of log lines graph}); this is explained fully in
\sectionref{Explaining the peaks in log file size}.

@@

\subsection{Explaining The Peaks In Log File Size}

\label{Explaining the peaks in log file size}

XXX SHOULD THIS BE A SEPARATE SECTION OR INCLUDED IN THE PREVIOUS ONE\@?
MAYBE INCLUDE GRAPHS OF RULES PER FILE OR SOMETHING TO EXTEND THIS\@?

XXX REWRITE WHEN STATS RUNS HAVE FINISHED\@; REWRITE ALL THE REFERENCES
TOO\@.

\Graphref{parsing time vs file size vs number of log lines graph} shows log
file size varying between 30--100MB, except for log files 22 (200MB) and
62--68 (400--900MB).  These large log files were caused by a user
unintentionally creating a mail forwarding loop, resulting in a huge number
of mail deliveries and up to an order of magnitude more log lines than is
usual.  The user specified his mail forwarding as: \newline{}
\tab{}\texttt{$\backslash$username, username@domain} \newline{} This
instructs Postfix to deliver the mail to the local user, and also forward a
copy to the remote address; this is generally not a problem except that the
remote address in this instance is the address the mail was originally sent
to, creating an infinite loop.  To prevent this happening Postfix examines
the Delivered-To header in the mail, and if the mail has already been
delivered to the current address it is bounced back to the sender with the
error message \newline{} \tab{} \texttt{mail forwarding loop for
username@domain}\newline{}  Normally this works well, but the forwarding
method in use at that time stripped the Delivered-To header from the mail,
breaking loop detection.

When the recipient in a mail loop is local, Postfix does not have to
deliver the mail to an external system, so mails are delivered as fast as
the disks can keep up, resulting in a huge volume of log lines.

XXX EXPAND THIS PARAGRAPH\@; THERE MAY BE CONTENT IN THE PAPER\@.  REFER TO
THE NEW GRAPH BELOW TOO\@.

The different distribution of log lines caused by a mail forwarding loop
has unexpected effects on parsing time, described in detail in
\sectionref{rule ordering for efficiency}, \sectionref{scalability as the
number of rules rises}, and \sectionref{Caching compiled regexes}.

\showgraph{build/graph-rules-tried-per-log-line}{Mean number of rules tried
per log line}{Mean number of rules tried per log line}

@@

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files (see \sectionref{Explaining the peaks in log file size}).
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of log lines vs.\ parsing time in
\graphref{parsing time vs file size vs number of log lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.

>>>>

XXX SHOULD I USE $\backslash{}$clearpage{} OR SOMETHING TO AFFECT FLOAT
PLACEMENT\@?  GRAPHS AND TABLES TEND TO MOVE QUITE FAR IN SOME OF THE
SUBSECTIONS, THOUGH MAKING THE TEXT LONGER MIGHT HELP TOO\@.  IDEALLY I
WANT TO HAVE TEXT, THEN ALL GRAPHS AND TABLES, THEN THE NEXT SECTION\@.

\newpage{} % XXX VARIOREF PROBLEMS

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse large log files.  The server that generated the log files used in
testing this parser accepts approximately 10,000 mails for 700 users a day;
\graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, far more mails are received on
weekdays than at weekends.  The mail server in question is a production
mail server handling mail for a university department; the benefit of using
this server is that its log files exhibit the idiosyncrasies and
peculiarities a mail server in the wild must deal with, but the downside is
that significantly altering the configuration to accommodate this project
is not an option.  Median log file size is 50MB, containing 285,000 log
lines; large scale mail servers would have much larger log files.  Note
that \graphref{Mails received per day} and \tableref{Number of mails
received per day: statistics} show the number of mails received by
\acronym{SMTP} only; in particular the mail loops noticeable in later
graphs do not contribute to these figures.

@@

@@

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, and the times from the 10 runs averaged.  The computer
used for test runs was a Dell Optiplex 745, shown in \tableref{Details of
the computer used to generate statistics}; it was dedicated to the task of
gathering statistics from test runs, and was not used for any other purpose
while test runs were ongoing.  Saving results to the database was disabled
for the test runs, because that dominates the run time of the parser, and
the tests are aimed at measuring the speed of \parsername{} rather than the
speed of the database and the disks the database is stored on.  Parsing all
\numberOFlogFILES{} log files in one run without saving results to the
database took \input{build/include-full-run-duration}, mean throughput is
\input{build/include-full-run-throughput}; median throughput when parsing
files separately was \input{build/include-median-throughput-MB}
(\input{build/include-median-throughput-log-lines}) parsed per minute.  In
contrast, when saving results to the database, parsing all
\numberOFlogFILES{} log files took
\input{build/include-insert-results-duration}, with mean throughput of
\input{build/include-insert-results-throughput}; parsing makes up
\input{build/include-skip-inserting-as-percentage-of-inserting} of the
execution time when saving results to the database, so a 50\% improvement
in the speed of parsing would result in only a
\input{build/include-skip-inserting-with-50-percent-improvement--as-percentage-of-inserting}
improvement in speed overall.

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{parsing time vs file size vs number of log lines graph} shows the
parsing time in seconds, file size in MB, and number of log lines in tens
of thousands, for each of the \numberOFlogFILES{} log files.  The three
lines run roughly in parallel, giving the impression that the algorithm
scales linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number of log lines factor}, which
plots both the ratio of file size vs parsing time, and the ratio of number
of log lines vs parsing time (higher is better in both cases);
\tableref{parsing time vs file size vs number of log lines factor table}
shows the same ratios for different groups of log files.  The ratios are
quite tightly banded, showing that the algorithm scales linearly; the
ratios increase (i.e.\ improve) for log files 22 and 62--68, despite their
larger than average size.  In both groups of log files there was a mail
loop because a user set up mail forwarding incorrectly, resulting in a very
different distribution of log lines: normally most log lines are generated
by mail delivery attempts from other hosts, but when the mail loops
occurred most of the log lines resulted from failed delivery of mail
generated on the server itself.  The Postfix components that generated most
of the log lines during the mail loop have fewer associated rules than the
Postfix components whose log lines normally make up the bulk of each log
file; \graphref{Average number of rules tried per log line} shows the drop
in the average number of rules tried per log line for log files containing
a mail loop.

@@

\showgraph{build/graph-average-number-of-rules-tried-per-log-line}{Average
number of rules tried per log line}{Average number of rules tried per log
line}

\showgraph{build/graph-average-number-of-rules-tried-per-program}{Average
number of rules tried per Postfix component per log line}{Average number of
rules tried per Postfix component per log line}

\FloatBarrier{}

@@

\newpage{} % XXX GRAPH PLACEMENT FLIP-FLOPPING

@@

Differences in rule ordering have less effect on parsing time when parsing
log files 22 and 62--68, due to the different distribution of log lines in
those log files.
When a mail loop occurs the vast majority of log lines are from Postfix
components that have few rules associated with them, whereas in general
most log lines are from Postfix components that have a high number of
associated rules; this results in the average number of rules required to
parse a log line from log files 22 or 62--68 being considerably lower than
usual.  The ratio of number of log lines vs.\ parsing time in
\graphref{parsing time vs file size vs number of log lines factor} shows a
noticeable improvement at log files 22 and 62--68, where a higher ratio
means more log lines parsed per unit of time; similarly \graphref{Parsing
time relative to shuffled ordering graph} shows a much smaller difference
between the orderings at log files 22 and 62--68.
